---
title: "Getting Started with PhyloBasins"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with PhyloBasins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

PhyloBasins is an R package for branch-based phylogenetic analyses of ecological communities. The package provides a reproducible workflow for validating phylogenetic trees, constructing branch tables, preparing community matrices, and generating sparse site × branch matrices that form the computational basis for phylogenetic diversity analyses.

This vignette introduces the core workflow of the package using a simple example.

## Workflow

```text
Phylogenetic tree
        │
        ▼
prepare_tree()

        │
        ▼
build_branch_table()

        │
Community matrix
        │
        ▼
prepare_community()

        │
        ▼
build_site_branch_matrix(pb)

        │
        ▼
Site × branch matrix
```

## Creating a project

The first step is to create a PhyloBasins project object.

```{r}
library(PhyloBasins)

pb <- pb_project()
```

The `pb_project()` constructor creates a project object that stores all input data and intermediate results throughout the analysis.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PhyloBasins)
```
